---
title: "Simluation plot for age of ischemic stroke onset."
author: "[Sander W. van der Laan, PhD](https://swvanderlaan.github.io) | @swvanderlaan | s.w.vanderlaan@gmail.com"
date: "`r Sys.Date()`"
output:
  html_notebook:
    cache: yes
    code_folding: hide
    collapse: yes
    df_print: paged
    fig.align: center
    fig_caption: yes
    fig_height: 6
    fig_retina: 2
    fig_width: 7
    highlight: tango
    theme: lumen
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes
mainfont: Arial
subtitle: "A 'druggable-MI-targets' project"
editor_options:
  chunk_output_type: inline
---

```{r global_options, include = FALSE}
# further define some knitr-options.
knitr::opts_chunk$set(fig.width = 12, fig.height = 8, fig.path = 'Figures/', 
                      wwarning = TRUE, # show warnings during codebook generation
  message = TRUE, # show messages during codebook generation
  error = TRUE, # do not interrupt codebook generation in case of errors,
                # usually better for debugging
  echo = TRUE,  # show R code
                      eval = TRUE)

```

# Setup

We will clean the environment, setup the locations, define colors, and create a datestamp.

*Clean the environment.*

```{r echo = FALSE}

rm(list = ls())

```

*Set locations and working directories...*

```{r LocalSystem, echo = FALSE}
### Operating System Version 
### MacBook Pro
ROOT_loc = "/Users/swvanderlaan/iCloud"
AEROOT_loc = "/Users/swvanderlaan"
# STORAGE_loc = "/Volumes/LaCie/"
STORAGE_loc = "/Users/swvanderlaan"

### MacBook Air
# ROOT_loc = "/Users/slaan3/OneDrive - UMC Utrecht"
# STORAGE_loc = "/Volumes/LaCie/"
# STORAGE_loc = "/Users/slaan3/"

GENOMIC_loc = paste0(ROOT_loc, "/Genomics")
AEDB_loc = paste0(AEROOT_loc, "/Athero-Express/AE-AAA_GS_DBs")
LAB_loc = paste0(GENOMIC_loc, "/LabBusiness")

PLINK_loc=paste0(STORAGE_loc,"/PLINK")
AEGSQC_loc =  paste0(PLINK_loc, "/_AE_ORIGINALS/AEGS_COMBINED_QC2018")
MICHIMP_loc=paste0(PLINK_loc,"/_AE_ORIGINALS/AEGS_COMBINED_EAGLE2_1000Gp3v5HRCr11")

GWAS_loc=paste0(PLINK_loc,"/_GWAS_Datasets/_SiGN")

PROJECT_loc = paste0(PLINK_loc, "/analyses/stroke/AAO_IschemicStroke")

# use this if there is relevant information here.
TARGET_loc = paste0(PROJECT_loc, "/targets")

### SOME VARIABLES WE NEED DOWN THE LINE
TRAIT_OF_INTEREST = "AAO" # Phenotype
PROJECTNAME = "AAO"

cat("\nCreate a new analysis directory...\n")
ifelse(!dir.exists(file.path(PROJECT_loc, "/",PROJECTNAME)), 
       dir.create(file.path(PROJECT_loc, "/",PROJECTNAME)), 
       FALSE)
ANALYSIS_loc = paste0(PROJECT_loc,"/",PROJECTNAME)

ifelse(!dir.exists(file.path(ANALYSIS_loc, "/PLOTS")), 
       dir.create(file.path(ANALYSIS_loc, "/PLOTS")), 
       FALSE)
PLOT_loc = paste0(ANALYSIS_loc,"/PLOTS")

ifelse(!dir.exists(file.path(PLOT_loc, "/QC")), 
       dir.create(file.path(PLOT_loc, "/QC")), 
       FALSE)
QC_loc = paste0(PLOT_loc,"/QC")

ifelse(!dir.exists(file.path(ANALYSIS_loc, "/OUTPUT")), 
       dir.create(file.path(ANALYSIS_loc, "/OUTPUT")), 
       FALSE)
OUT_loc = paste0(ANALYSIS_loc, "/OUTPUT")

ifelse(!dir.exists(file.path(ANALYSIS_loc, "/BASELINE")), 
       dir.create(file.path(ANALYSIS_loc, "/BASELINE")), 
       FALSE)
BASELINE_loc = paste0(ANALYSIS_loc, "/BASELINE")

setwd(paste0(PROJECT_loc))
getwd()
list.files()

```

*... a package-installation function ...*

```{r}
source(paste0(PROJECT_loc, "/scripts/functions.R"))

```

*... and load those packages.*

```{r loading_packages, message=FALSE, warning=FALSE}
source(paste0(PROJECT_loc, "/scripts/packages.R"))

```

```{r}
ggplot2::theme_set(ggplot2::theme_minimal())
pander::panderOptions("table.split.table", Inf)
```

_We will create a datestamp and define the Utrecht Science Park Colour Scheme_.
```{r Setting: Colors}

Today = format(as.Date(as.POSIXlt(Sys.time())), "%Y%m%d")
Today.Report = format(as.Date(as.POSIXlt(Sys.time())), "%A, %B %d, %Y")

### UtrechtScienceParkColoursScheme
###
### WebsitetoconvertHEXtoRGB:http://hex.colorrrs.com.
### Forsomefunctionsyoushoulddividethesenumbersby255.
###
###	No.	Color			      HEX	(RGB)						              CHR		  MAF/INFO
###---------------------------------------------------------------------------------------
###	1	  yellow			    #FBB820 (251,184,32)				      =>	1		or 1.0>INFO
###	2	  gold			      #F59D10 (245,157,16)				      =>	2		
###	3	  salmon			    #E55738 (229,87,56)				      =>	3		or 0.05<MAF<0.2 or 0.4<INFO<0.6
###	4	  darkpink		    #DB003F ((219,0,63)				      =>	4		
###	5	  lightpink		    #E35493 (227,84,147)				      =>	5		or 0.8<INFO<1.0
###	6	  pink			      #D5267B (213,38,123)				      =>	6		
###	7	  hardpink		    #CC0071 (204,0,113)				      =>	7		
###	8	  lightpurple	    #A8448A (168,68,138)				      =>	8		
###	9	  purple			    #9A3480 (154,52,128)				      =>	9		
###	10	lavendel		    #8D5B9A (141,91,154)				      =>	10		
###	11	bluepurple		  #705296 (112,82,150)				      =>	11		
###	12	purpleblue		  #686AA9 (104,106,169)			      =>	12		
###	13	lightpurpleblue	#6173AD (97,115,173/101,120,180)	=>	13		
###	14	seablue			    #4C81BF (76,129,191)				      =>	14		
###	15	skyblue			    #2F8BC9 (47,139,201)				      =>	15		
###	16	azurblue		    #1290D9 (18,144,217)				      =>	16		or 0.01<MAF<0.05 or 0.2<INFO<0.4
###	17	lightazurblue	  #1396D8 (19,150,216)				      =>	17		
###	18	greenblue		    #15A6C1 (21,166,193)				      =>	18		
###	19	seaweedgreen	  #5EB17F (94,177,127)				      =>	19		
###	20	yellowgreen		  #86B833 (134,184,51)				      =>	20		
###	21	lightmossgreen	#C5D220 (197,210,32)				      =>	21		
###	22	mossgreen		    #9FC228 (159,194,40)				      =>	22		or MAF>0.20 or 0.6<INFO<0.8
###	23	lightgreen	  	#78B113 (120,177,19)				      =>	23/X
###	24	green			      #49A01D (73,160,29)				      =>	24/Y
###	25	grey			      #595A5C (89,90,92)				        =>	25/XY	or MAF<0.01 or 0.0<INFO<0.2
###	26	lightgrey		    #A2A3A4	(162,163,164)			      =>	26/MT
### 
###	ADDITIONAL COLORS
###	27	midgrey			#D7D8D7
###	28	verylightgrey	#ECECEC"
###	29	white			#FFFFFF
###	30	black			#000000
###----------------------------------------------------------------------------------------------

uithof_color = c("#FBB820","#F59D10","#E55738","#DB003F","#E35493","#D5267B",
                 "#CC0071","#A8448A","#9A3480","#8D5B9A","#705296","#686AA9",
                 "#6173AD","#4C81BF","#2F8BC9","#1290D9","#1396D8","#15A6C1",
                 "#5EB17F","#86B833","#C5D220","#9FC228","#78B113","#49A01D",
                 "#595A5C","#A2A3A4", "#D7D8D7", "#ECECEC", "#FFFFFF", "#000000")

uithof_color_legend = c("#FBB820", "#F59D10", "#E55738", "#DB003F", "#E35493",
                        "#D5267B", "#CC0071", "#A8448A", "#9A3480", "#8D5B9A",
                        "#705296", "#686AA9", "#6173AD", "#4C81BF", "#2F8BC9",
                        "#1290D9", "#1396D8", "#15A6C1", "#5EB17F", "#86B833",
                        "#C5D220", "#9FC228", "#78B113", "#49A01D", "#595A5C",
                        "#A2A3A4", "#D7D8D7", "#ECECEC", "#FFFFFF", "#000000")
### ----------------------------------------------------------------------------
```

# Introduction

We will parse the data to create regional association plots for each of the 11 loci. 

# Setting the NPG colors

```{r}
library("scales")
pal_npg("nrc")(10)
show_col(pal_npg("nrc")(10))

# show_col(pal_npg("nrc", alpha = 0.6)(10))

```


# Load data

We need to load the data first. This was generated in SAS with `scripts/simulateCaseControl.sas`.

```{r}

simdata <- fread(file = paste0(PROJECT_loc, "/simulation_data/simulation_data.csv"))

# str(simulation_data)
# Classes ‘data.table’ and 'data.frame':	33 obs. of  5 variables:
#  $ study  : chr  "Additive SNP Effect on Stroke" "Multiplicative SNP Effect on Stroke" "SNP Effect on Death" "Additive SNP Effect on Stroke" ...
#  $ gEffect: num  1 1 1 1.1 1.1 1.1 1.2 1.2 1.2 1.3 ...
#  $ reject : int  1 1 1 1 1 1 1 1 1 1 ...
#  $ COUNT  : int  41 48 55 50 64 57 102 111 48 180 ...
#  $ PERCENT: num  4.1 4.8 5.5 5 6.4 5.7 10.2 11.1 4.8 18 ...
#  - attr(*, ".internal.selfref")=<externalptr>
#   
```

```{r}
library(dplyr)
simdata <- simdata %>%
    mutate(study = recode(study, 
                           "Additive SNP Effect on Stroke" = "Additive SNP effect on stroke", 
                           "Multiplicative SNP Effect on Stroke" = "Multiplicative SNP effect on stroke", 
                           "SNP Effect on Death" = "SNP effect on death"))
```


# Simulation plotting

```{r}

ggpubr::ggline(simdata,
               x = "gEffect",
               y = "PERCENT",
               color = "study", palette = "npg",
               plot_type = "l", 
               linetype = 1,
               size = 1.25,
               # shape = 19, 
               # stroke = 2,
               point.size = 1.25,
               # point.color = "npg",
               xlab = "Genetic effect size", ylab = "Power",
               ylim = c(0, 100),
               legend = "right",
               ggtheme = theme_minimal()) +
  theme(axis.line = element_line(size = 1, colour = uithof_color[25]),
        legend.position = c(0, 1),
        legend.justification = c("left", "top"),
        legend.box.just = "left",
        legend.margin = margin(6, 6, 6, 6)) +
  labs(color = "Simulated genetic effect")

ggsave(paste0(PROJECT_loc, "/simulation_graph/SimulationResults.png"), plot = last_plot())
ggsave(paste0(PROJECT_loc, "/simulation_graph/SimulationResults.pdf"), plot = last_plot())
ggsave(paste0(PROJECT_loc, "/simulation_graph/SimulationResults.eps"), plot = last_plot())

```


# Session information

------

    Version:      v1.0.0
    Last update:  2022-05-04
    Written by:   Sander W. van der Laan (s.w.vanderlaan-2[at]umcutrecht.nl).
    Description:  Script to create simulation plot.
    Minimum requirements: R version 3.4.3 (2017-06-30) -- 'Single Candle', Mac OS X El Capitan
    
    Changes log
    * v1.0.0 Initial version. 

------

```{r eval = TRUE}
sessionInfo()
```


# Saving environment
```{r Saving}

save.image(paste0(PROJECT_loc, "/",Today,".",PROJECTNAME,".SimulationPlot.RData"))
```


------
<sup>&copy; 1979-2022 Sander W. van der Laan | s.w.vanderlaan[at]gmail.com | [swvanderlaan.github.io](https://swvanderlaan.github.io).</sup>
------

  
